A Constraint Solving Approach to Parikh Images of Regular Languages

### 2023-07-11 multiple.tex
Math!cal!

We return to our example in \cref{sec:intuition}. We also reuse the
definition of $\Map$ as the simple projection to extract the counter increments,
so in this case, $\Map(\Transition) = \left[l_{B+}\right]$ for every
transition~$\Transition$ since $B$ only counts length. The reader is
strongly advised to review \cref{fig:aut_a_annotated,fig:aut_b_annotated} from
\cref{sec:intuition} while going through this example! In particular, note that
the transition variables of the figures match the ones used here.

We extend \cref{eq:single-filter} from \cref{sec:single:example} with
definitions for $B$ shown in \cref{eq:multi-filter}:
\begin{equation}\label{eq:multi-filter}
    \begin{aligned}
        & \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) = \TransitionVar_6 
        & \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7 
        & \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B}) = \TransitionVar_8 
        &\Filter(\FromLabelTo{B}{c / l_{B+}}{B}) = \TransitionVar_9  \\
        & \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B})= \TransitionVar_{10}
        & \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A})= \TransitionVar_{11}
        & \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}
        \FlowEq(B, \Filter) &=  \begin{aligned}
            1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
            \land \TransitionVar_{11} &= \TransitionVar_6\\
            \land \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}
        \end{aligned}
    \end{aligned}
\end{equation}

@@

We then continue by repeating the steps of \cref{sec:single:example} shown in
\cref{fig:derivation:single}, since the steps to propagate
\cref{ex:string-constraints} across the automaton $A'$ still apply for the
product. Note that this fixes the values of all transitions but one of $A'$ as
constants, including the free counter variables $i, n$, whose new constant
values we propagate everywhere.

Once we have performed reasoning by equality to eliminate transition variables,
we obtain the version of $B$ shown in \cref{fig:aut_b_annotated} and represented
by the clauses in the final node in the tree. This corresponds to the end of
\cref{sec:intuition:algebra}. Our only options now are to either split on which
path we take from $I_B$ by case-splitting on~$\TransitionVar_{12}$ or to
directly invoke \Materialise{}. The latter would produce a shorter tree in this
instance but might lead to a larger product being computed, so we (and indeed
\Catra{}) pick \Split{}. This leads to an opportunity for propagation since
$\TransitionVar_{12} = 0$ cuts off $\TransitionVar_{9}$ from $I_B$ on the left
branch. Note that the same opportunity does not present itself on the right
branch, though; the use of the upper path to state~$A$ does not preclude visiting
state~$B$.

@@

      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A})

@@

      \land \Connected(B, \Filter) \\
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 

@@

      \Connected(B, \Filter) \land\\
      \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 

@@

    \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 

@@

      \land \Connected(B, \Filter) 
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 

@@

      \land \FlowEq(B, \Filter) \\
      \land \Connected(B, \Filter) 
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 

@@

      \land \FlowEq(B, \Filter)
      \land \Connected(A', \Filter)\\
      \land \Connected(B, \Filter)
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B}) \\

@@

    \Image{}_{\Tuple{A', B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B})

>>>>

We return to our example in \cref{sec:intuition}. We also reuse the
definition of $\Map$ as the simple projection to extract the counter increments,
so in this case, $\Map(\Transition) = \left[l_{B+}\right]$ for every
transition~$\Transition$ since $\mathcal{B}$ only counts length. The reader is
strongly advised to review \cref{fig:aut_a_annotated,fig:aut_b_annotated} from
\cref{sec:intuition} while going through this example! In particular, note that
the transition variables of the figures match the ones used here.

We extend \cref{eq:single-filter} from \cref{sec:single:example} with
definitions for $B$ shown in \cref{eq:multi-filter}:
\begin{equation}\label{eq:multi-filter}
    \begin{aligned}
        & \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) = \TransitionVar_6 
        & \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7 
        & \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B}) = \TransitionVar_8 
        &\Filter(\FromLabelTo{B}{c / l_{B+}}{B}) = \TransitionVar_9  \\
        & \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B})= \TransitionVar_{10}
        & \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A})= \TransitionVar_{11}
        & \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}
        \FlowEq(\mathcal{B}, \Filter) &=  \begin{aligned}
            1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
            \land \TransitionVar_{11} &= \TransitionVar_6\\
            \land \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}
        \end{aligned}
    \end{aligned}
\end{equation}

@@

We then continue by repeating the steps of \cref{sec:single:example} shown in
\cref{fig:derivation:single}, since the steps to propagate
\cref{ex:string-constraints} across the automaton $\mathcal{A}'$ still apply for the
product. Note that this fixes the values of all transitions but one of $\mathcal{A}'$ as
constants, including the free counter variables $i, n$, whose new constant
values we propagate everywhere.

Once we have performed reasoning by equality to eliminate transition variables,
we obtain the version of $\mathcal{B}$ shown in \cref{fig:aut_b_annotated} and represented
by the clauses in the final node in the tree. This corresponds to the end of
\cref{sec:intuition:algebra}. Our only options now are to either split on which
path we take from $I_B$ by case-splitting on~$\TransitionVar_{12}$ or to
directly invoke \Materialise{}. The latter would produce a shorter tree in this
instance but might lead to a larger product being computed, so we (and indeed
\Catra{}) pick \Split{}. This leads to an opportunity for propagation since
$\TransitionVar_{12} = 0$ cuts off $\TransitionVar_{9}$ from $I_B$ on the left
branch. Note that the same opportunity does not present itself on the right
branch, though; the use of the upper path to state~$A$ does not preclude visiting
state~$B$.

@@

      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A})

@@

      \land \Connected(\mathcal{B}, \Filter) \\
      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 

@@

      \Connected(\mathcal{B}, \Filter) \land\\
      \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 

@@

    \land \Connected(\mathcal{B}, \Filter)\\
    \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 

@@

      \land \Connected(\mathcal{B}, \Filter) 
      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 

@@

      \land \FlowEq(\mathcal{B}, \Filter) \\
      \land \Connected(\mathcal{B}, \Filter) 
      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 

@@

      \land \FlowEq(\mathcal{B}, \Filter)
      \land \Connected(\mathcal{A}', \Filter)\\
      \land \Connected(\mathcal{B}, \Filter)
      \land \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B}) \\

@@

    \Image{}_{\Tuple{\mathcal{A}',\mathcal{B}}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B})
